Swinging, Fast and Slow

### 2024-08-24 reports/saberseminar/slides.tex
final? commit before saberseminar

\begin{frame}{Adaptation: How do intended swings vary by location?}
  \centering
  \includegraphics[width = \textwidth]{../../figures/beamer/adaptation.pdf}
\end{frame}

\begin{frame}{Timing: What swing variation remains unexplained?}
  \includegraphics[width = 0.49\textwidth]{../../figures/beamer/swing_metrics_intended.pdf}
  \includegraphics[width = 0.49\textwidth]{../../figures/beamer/swing_metrics_residual.pdf}
\end{frame}

@@

\begin{frame}{Approach Model}
  \begin{enumerate}
    \item Start with pitch outcome model which estimates $\mathbb{P}$(swing), $\mathbb{P}$(contact $\mid$ swing), $\mathbb{P}$(fair $\mid$ contact), $\mathbb{E}$(xwOBA $\mid$ fair), etc.
    \begin{itemize}
      \item Gradient boosting w/ 2022-2024 MLB data
    \end{itemize}
    \item Refit contact, fair and xwOBA models on batter approach, with previous predictions as offset
  \end{enumerate}
  Example: $p_i$ is contact prob, $\hat p_i$ is pitch outcome model estimate
  \begin{align*}
    \log\left(\frac{p_i}{1 - p_i}\right) = \log\left(\frac{\hat p_i}{1 - \hat p_i}\right) + \beta \cdot \hat\gamma^S_{b_i} \cdot (\mbox{strikes})_i
  \end{align*}
  \begin{itemize}
    \item This is instrumental variable regression from causal inference, using count as the instrument
  \end{itemize}
\end{frame}

\begin{frame}{The Effect of Approach}
\end{frame}

>>>>

\begin{frame}{Adaptation: How do intended swings vary by location?}
  \centering
  \includegraphics[width = \textwidth]{../../figures/beamer/adaptation.pdf}
  \begin{itemize}
    \item There's clearly a difference, but it's hard to say why
  \end{itemize}
\end{frame}

\begin{frame}{Timing: What swing variation remains unexplained?}
  \includegraphics[width = 0.49\textwidth]{../../figures/beamer/swing_metrics_intended.pdf}
  \includegraphics[width = 0.49\textwidth]{../../figures/beamer/swing_metrics_residual.pdf}
  \begin{itemize}
    \item Left: Predicted intention for all swings (not just squared up)
    \item Right: Residuals (relative to intention) for all swings
    \begin{itemize}
      \item Pattern matches bat speed through path of swing!
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Approach Model}
  \begin{enumerate}
    \item Start with pitch outcome model which estimates $\mathbb{P}$(swing), $\mathbb{P}$(contact $\mid$ swing), $\mathbb{P}$(fair $\mid$ contact), $\mathbb{E}$(xwOBA $\mid$ fair), etc.
    \begin{itemize}
      \item Gradient boosting w/ 2022-2024 MLB data
    \end{itemize}
    \item Refit contact, fair and xwOBA models on batter approach, with previous predictions as offset
  \end{enumerate}
  Example: $p_i$ is contact prob, $\hat p_i$ is pitch outcome model estimate
  \begin{align*}
    \log\left(\frac{p_i}{1 - p_i}\right) = \log\left(\frac{\hat p_i}{1 - \hat p_i}\right)
      &+ \beta^{SL} \cdot \hat\gamma^{SL}_{b_i} \cdot (\mbox{strikes})_i\\
      &+ \beta^{BS} \cdot \hat\gamma^{BS}_{b_i} \cdot (\mbox{strikes})_i
  \end{align*}
  \begin{itemize}
    \item This is instrumental variable regression from causal inference, using count as the instrument
  \end{itemize}
\end{frame}

\begin{frame}{The Effect of Approach}{``Effect'' is approach-aware vs. approach-agnostic prediction w/ 2 strikes}
  \centering
  \includegraphics[width = 0.49\textwidth]{../../figures/beamer/swing_length_contact.pdf}
  \includegraphics[width = 0.49\textwidth]{../../figures/beamer/bat_speed_contact.pdf}\\
  \includegraphics[width = 0.49\textwidth]{../../figures/beamer/swing_length_power.pdf}
  \includegraphics[width = 0.49\textwidth]{../../figures/beamer/bat_speed_power.pdf}
\end{frame}
